Reducing retrieval error in MAIAC AOD with the `Earth_obs_cleaning` machine-learning workflow: Results from CONUS

### 2023-08-16 writing/CONUS_AOD.qmd
Removed an obsolete comment.
Deduplicated the code for agreement plots.

# Contiguous US state boundaries
# will replace with a region defined by Kodi later
conus = st_transform(crs = crs.us.atlas, get_conus())
```

@@

The original error envelope (reported in the AMT MAIAC publication) had `r dt[, round(100 * mean(y.sat > (y.ground - 0.05 - 0.1*y.ground) & y.sat < (y.ground + 0.05 + 0.1*y.ground)))][1]`% coverage in our data.

```{r error_envelope}
# now with our error envelope function
envelope.sat <- empirical.error.envelope(dt, pred = "y.sat")
kable(head(envelope.sat, 1), digits = 4)

# and for the corrected data
envelope.ground.pred <- empirical.error.envelope(dt, pred = "y.ground.pred")
kable(head(envelope.ground.pred, 1), digits = 4)

```

@@

```{r agreement_plot, dev='png', dpi=300}
p1 <- ggplot(dt[y.ground < 1 & y.sat < 1]) +
  aes(y.ground, y.sat) +
  geom_pointdensity(alpha = 0.1, size = 0.6) +
  geom_abline(linetype = "dashed") +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
  guides(color = "none") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  labs(caption = paste0("Not shown: ", dt[y.ground > 1 | y.sat > 1, .N], " AOD values outside [0, 1]"))

p2  <- ggplot(dt[y.ground < 1 & y.ground.pred > 0 & y.ground.pred < 1]) +
  aes(y.ground, y.ground.pred) +
  geom_pointdensity(alpha = 0.1, size = 0.6) +
  geom_abline(linetype = "dashed") +
  geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
  scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
  scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
  guides(color = "none") +
  theme_bw() +
  theme(aspect.ratio = 1) +
  labs(caption = paste0("Not shown: ", dt[y.ground > 1 | y.ground.pred > 1 | y.ground.pred < 0, .N], " AOD values outside [0, 1]"))

p1 + p2
```

>>>>

# Contiguous US state boundaries
conus = st_transform(crs = crs.us.atlas, get_conus())
```

@@

We use Nelder-Mead optimization to construct new error envelopes with coverage as close to 2/3 as possible for each of two subsets of the data: observations less than or equal to 0.6, and observations greater than 0.6. The results are an envelope with `r f = \(pred) {ps = empirical.error.envelope(obs = dt$y.ground, pred)$par; sprintf("additive part %.03f and multiplicative part %.03f", ps[1], ps[2])}; f(dt$y.sat)` for the raw satellite values, and an envelope with `r f(dt$y.ground.pred)` for our predictions. The original error envelope (reported in the AMT MAIAC publication), applied to the raw satellite values, has coverage `r x = envelope.tester(obs = dt$y.ground, pred = dt$y.sat)(.05, .1); sprintf("%.03f", x["lo"])` in the low subset and `r sprintf("%.03f", x["hi"])` in the high subset.

@@

```{r agreement_plot, dev='png', dpi=300}
agreement.plot = \(vname.pred)
   {i = dt[, between(y.ground, 0, 1) & between(get(vname.pred), 0, 1)]
    ggplot(dt[i]) +
        aes(y.ground, get(vname.pred)) +
        geom_pointdensity(alpha = 0.1, size = 0.6) +
        geom_abline(linetype = "dashed") +
        geom_smooth(method = "gam", formula = y ~ s(x, bs = "cs")) +
        scale_x_continuous("AERONET AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
        scale_y_continuous("MAIAC AOD 470nm", limits = c(0, 1), expand = c(0, 0)) +
        guides(color = "none") +
        theme_bw() +
        theme(aspect.ratio = 1) +
        labs(caption = paste0("Not shown: ", dt[!i, .N], " AOD values outside [0, 1]"))}

agreement.plot("y.sat") + agreement.plot("y.ground.pred")
```
